---
title: "Data + statistical analyses and figures for Larsen et al. 2019"
subtitle: "Conestogo Lake, Mapleton Township, Ontario"
output: 
  pdf_document:
    latex_engine: xelatex
    citation_package: natbib
geometry: left = 2cm, right = 2cm, top = 2cm, bottom = 2cm
fontsize: 12pt
mainfont: HelveticaNeue Light
bibliography: reportcitations
editor_options: 
  chunk_output_type: console
---

```{r, include = F}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      fig.height = 5.5)

#, fig.pos = 'H'
#results = "hide"

# libraries
library(tidyverse)
library(akima)
library(lubridate)
library(padr)
library(gridExtra)
library(pander)
library(RCurl)
library(MBA)
library(reshape)
library(viridis)
library(scales)
library(cowplot)
library(nlme)
library(lemon)
library('vegan'); library(ade4)
library(lsmeans)


# Plotting theme
library(extrafont)
font_import()
#loadfonts(device="win")

# load functions
source("../database_Conestogo/src/themeformatting.R")
source("../database_Conestogo/src/sondeinterpolationplot.R")
source("../database_Conestogo/src/cheminterpolationplot.R")

# set theme
theme_set(theme_std(base_family = "", base_size = 11))
#ggplot(mtcars, aes(x = cyl, y =mpg)) + geom_point() + theme_std()
# tables shoud have decimal alignment
```

\vspace*{5cm}
M.L. Larsen^1^, H. Baulch^2^, S. Schiff^3^,Dana F. Simon^4^, Sesbastien Sauve^4^, J. Venkiteswaran^1^

\vspace*{5cm}
^1^Department of Geography and Environmental Studies, Wilfrid Laurier University
^2^School of Environment and Sustainability, Global Institute for Water Security, University of Saskatchewan 
^3^Department of Earth and Environmental Sciences, University of Waterloo 
^4^Department of Chemistry, Universite de Montreal

\vspace*{3cm}
Citation: 


```{r Versions}
cap.text = paste("This analysis was produced using", R.version.string, ".")

```

# Time-series statistics

```{r chemdataImport}
chem.dat <- read.csv("./data/con_chemistry_profiles.csv", header = T) %>%
  mutate(Date = as.Date(Date))

res.dat <- read.csv("./data/reservoirhistorical.csv", header = T)

res.dat <- res.dat %>%
  mutate(date = as.Date(date)) %>%
  filter(date >= as.Date("2017-06-01"), date <= as.Date("2017-08-17"))

#merge.chem.dat <- chem.dat %>%
#  left_join(con.dat, by = c("Date" = "Date", "Site" = "Site")) %>%
#  left_join(res.levels.daily, by = c("Date" = "date")) %>%
#  mutate(bottom.elev = mean.value - Depth_M,
#         ele.depth = mean.value - Depth, DOY = yday(Date)) %>%
#  select(Site, Date, DOY, Depth:TP, Secchi, Air_Temp:mean.value, ele.depth)

merge.chem.dat <- chem.dat %>%
  left_join(res.dat, by = c("Date" = "date")) %>%
  mutate(ele.depth = elevation.m_mean - Depth, DOY = yday(Date),
         station = Site) 

#head(chem.dat)

chem.dat.epi <- chem.dat %>%
  filter(Depth == 2) %>%
  mutate(Date = as.Date(Date, format = "%y-%m-%d")) %>%
  dplyr::select(Site:Date, Fe:TDN, DOC, TDP:TP)

chem.dat.meta <- chem.dat %>%
  filter(Depth <= 7, Depth != 2) %>%
  mutate(Date = as.Date(Date, format = "%y-%m-%d")) %>%
  dplyr::select(Site:Date, Fe:TDN, DOC, TDP:TP)

chem.dat.hypo <- chem.dat %>%
  filter(Depth != 7, Depth != 2) %>%
  mutate(Date = as.Date(Date, format = "%y-%m-%d")) %>%
  dplyr::select(Site:Date, Depth, Fe:TDN, DOC, TDP:TP)

```


```{r chem-RMANOVA-data}
p.list <- c("SRP","TDP","TP", "Mn")
pars.list <- c("DOC","NH4","NO2","NO3","TDN","Cl","SO4","Mg","Ca","Fe")
epi.chem <- chem.dat %>%
  mutate(DOY = yday(Date)) %>%
  filter(Site != "clw", Depth == 2, DOY > 180) %>%
  dplyr::select(Site, Date, DOY, p.list, pars.list, -NH4) 
```

## SRP
```{r SRP-RMANOVA}
# test model with random effects vs. just fixed effects
model <- lme(data = epi.chem, 
           SRP ~ DOY * Site,  
           random = ~1|Site, 
           correlation = corAR1(form = ~ DOY|Site), 
           method = "REML", na.action=na.omit)

model.fixed <- gls(SRP ~ DOY * Site,
    correlation = corAR1(form = ~ DOY|Site),
    data = epi.chem,
    method = "REML")

anova(model, model.fixed) ## There is no benefit to having site as a random variable.

```

```{r SRP-RMANOVA-table}
pander(anova(model.fixed))
```

## TDP
```{r TDP-RMANOVA}
# test model with random effects vs. just fixed effects
model <- lme(data = epi.chem, 
           TDP ~ DOY * Site,  
           random = ~1|Site, 
           correlation = corAR1(form = ~ DOY|Site), 
           method = "REML", na.action=na.omit)

model.fixed <- gls(TDP ~ DOY * Site,
    correlation = corAR1(form = ~ DOY|Site),
    data = epi.chem,
    method = "REML")

anova(model, model.fixed) ## There is no benefit to having site as a random variable.

```

```{r TDP-RMANOVA-table}
pander(anova(model.fixed))
```

## NO3
```{r NO3-RMANOVA}

# test model with random effects vs. just fixed effects
model <- lme(data = epi.chem, 
           NO3 ~ DOY * Site,  
           random = ~1|Site, 
           correlation = corAR1(form = ~ DOY|Site), 
           method = "REML", na.action=na.omit)

model.fixed <- gls(NO3 ~ DOY * Site,
    correlation = corAR1(form = ~ DOY|Site),
    data = epi.chem,
    method = "REML")

anova(model, model.fixed) ## There is no benefit to having site as a random variable.
```

```{r NO3-RMANOVA-table}
pander(anova(model.fixed))
```

## TDN
```{r TDN-RMANOVA}
# test model with random effects vs. just fixed effects
model <- lme(data = epi.chem, 
           TDN ~ DOY * Site,  
           random = ~1|Site, 
           correlation = corAR1(form = ~ DOY|Site), 
           method = "REML", na.action=na.omit)

model.fixed <- gls(TDN ~ DOY * Site,
    correlation = corAR1(form = ~ DOY|Site),
    data = epi.chem,
    method = "REML")

anova(model, model.fixed) ## There is no benefit to having site as a random variable.

```

```{r TDN-RMANOVA-table}
pander(anova(model.fixed))
```


## Chla

```{r ttest}
sonde.dat <- read.csv("./data/Conestogo_SondeProfiles_2017_processed.csv")

epi.sonde <- sonde.dat %>%
  filter(depth.adj.new == 2, DOY > 180)
```

```{r chla-RMANOVA}
# test model with random effects vs. just fixed effects
model <- lme(data = epi.sonde, 
           chla_ug.L_avg ~ DOY * station,  
           random = ~1|station, 
           correlation = corAR1(form = ~ DOY|station), 
           method = "REML", na.action=na.omit)

model.fixed <- gls(chla_ug.L_avg ~ DOY * station,
    correlation = corAR1(form = ~ DOY|station),
    data = epi.sonde,
    method = "REML")

anova(model, model.fixed) ## There is no benefit to having site as a random variable.
```

```{r chla-RMANOVA-table}
pander(anova(model.fixed))
```

## Dissolved Oxygen
```{r DO-RMANOVA}
# test model with random effects vs. just fixed effects
model <- lme(data = epi.sonde, 
           DO_mg.L_avg ~ DOY * station,  
           random = ~1|station, 
           correlation = corAR1(form = ~ DOY|station), 
           method = "REML", na.action=na.omit)

model.fixed <- gls(DO_mg.L_avg ~ DOY * station,
    correlation = corAR1(form = ~ DOY|station),
    data = epi.sonde,
    method = "REML")

anova(model, model.fixed) ## There is no benefit to having site as a random variable.

```

```{r DO-RMANOVA-table}
pander(anova(model.fixed))
```


## Temperature
```{r temp-RMANOVA}
# test model with random effects vs. just fixed effects
model <- lme(data = epi.sonde, 
           temp_C_avg ~ DOY * station,  
           random = ~1|station, 
           correlation = corAR1(form = ~ DOY|station), 
           method = "REML", na.action=na.omit)

model.fixed <- gls(temp_C_avg ~ DOY * station,
    correlation = corAR1(form = ~ DOY|station),
    data = epi.sonde,
    method = "REML")

anova(model, model.fixed) ## There is no benefit to having site as a random variable.
```

```{r temp-RMANOVA-table}
pander(anova(model.fixed))
```

## PAR 
```{r}
par.data <- read.csv("./data/par_merged.csv",header = T) %>%
  mutate(date = as.Date(date, format = "%Y-%m-%d"),
         DOY = yday(date)) %>%
  filter(Depth == 2, site != "clw", DOY > 180) %>%
  dplyr::select(site, date, DOY, PAR)

```

```{r PAR-RMANOVA}
# test model with random effects vs. just fixed effects
model <- lme(data = par.data, 
           PAR ~ DOY * site,  
           random = ~1|site, 
           correlation = corAR1(form = ~ DOY|site), 
           method = "REML", na.action=na.omit)

model.fixed <- gls(PAR ~ DOY * site,
    correlation = corAR1(form = ~ DOY|site),
    data = par.data,
    method = "REML")

anova(model, model.fixed) ## There is no benefit to having site as a random variable.
anova(model.fixed)

```

```{r PAR-RMANOVA-table}
pander(anova(model.fixed))
```


## Table 1
```{r}
epi.sonde <- sonde.dat %>%
  filter(depth.adj.new == 2, station != "clw", DOY > 180) %>%
  group_by(station) %>%
  summarize_at(vars(temp_C_avg:chla_ug.L_avg), list(mean = mean, min = min, max = max)) %>%
  gather(key = parameter, value = result, temp_C_avg_mean:chla_ug.L_avg_max) %>%
  separate(parameter, into = c("variable","metric"), sep = "_avg_") %>%
  spread(key = metric, value = result)

epi.chem <- chem.dat %>%
  mutate(DOY = yday(Date))%>%
  filter(Site != "clw", Depth == 2, DOY > 180) %>%
  dplyr::select(Site, Date, p.list, pars.list, -NH4) %>%
  group_by(Site) %>%
  summarize_at(vars(SRP:Ca), list(mean = mean, min = min, max = max), na.rm = TRUE) %>%
  gather(key = parameter, value = result, SRP_mean:Ca_max) %>%
  separate(parameter, into = c("variable","metric"), sep = "_") %>%
  spread(key = metric, value = result)


m <- data.frame(epi.sonde[1:4,2],
      paste0(round(epi.sonde$mean[5:8], digits = 2)," (", 
             round(epi.sonde$min[5:8], digits = 2)," - ",
             round(epi.sonde$max[5:8], digits = 2),")"),
      paste0(round(epi.sonde$mean[1:4], digits = 2)," (", 
             round(epi.sonde$min[1:4], digits = 2)," - ",
             round(epi.sonde$max[1:4], digits = 2),")"))
m2 <- data.frame(epi.chem$variable[1:12],
                 paste0(round(epi.chem$mean[1:12], digits = 2)," (", 
                        round(epi.chem$min[1:12], digits = 2), " - ", 
                        round(epi.chem$max[1:12], digits = 2), ")"),
                 paste0(round(epi.chem$mean[13:24], digits = 2)," (", 
                        round(epi.chem$min[13:24], digits = 2), " - ", 
                        round(epi.chem$max[13:24], digits = 2) , ")"))

colnames(m) <- c("Parameter","Transitional", "Lacustrine")
colnames(m2) <- c("Parameter","Transitional", "Lacustrine")

m3 <- t(bind_rows(m, m2))
colnames(m3) <- m3[1,]
m3 <- m3[-1,]

pander(t(m3))
```


### Phytoplankton community ordination

<Environmental Matrix>

```{r}

con.dat <- read.csv("./data/con_met_data.csv", 
                    header = T) %>% 
  mutate(Date = as.Date(Date, format = "%d-%b-%y")) %>%
  dplyr::select(Site:Secchi)

#head(con.dat)

res.dat <- read.csv("./data/reservoirhistorical.csv", 
                    header = T) %>%
  mutate(date = as.Date(date, format = "%Y-%m-%d")) %>%
  dplyr::select(date:wind.speed.km.hr_mean,precip.mm_sum, res.time)

# Create merged data file
res.dat.merged <- con.dat %>%
  left_join(res.dat, by = c("Date" = "date")) 
```

<EM: Sonde data| temp, pH, DO>
```{r SondeDataImport}

# sonde data for each site and date
sonde.dat <- read.csv("./data/Conestogo_SondeProfiles_2017_processed.csv", header = T)

# create a epi dataset
sonde.epi <- sonde.dat %>%
  filter(depth.adj.new == 2) %>%
  dplyr::select(station:depth.adj.new, temp_C_avg:pH_avg) %>%
  gather(key = parameter, value = result, temp_C_avg:pH_avg) %>%
  mutate(parameter = paste0(parameter,"_epi")) 

sonde.sed <- sonde.dat %>% 
  group_by(station, date) %>% 
  slice(n()) %>%
  dplyr::select(station:depth.adj.new, temp_C_avg:pH_avg) %>%
  gather(key = parameter, value = result, temp_C_avg:pH_avg) %>%
  mutate(parameter = paste0(parameter,"_sed"))

# lake physical characteristics at each site
lake.phys <- read.csv("./data/Conestogo_WaterColumnMetrics_2017.csv", header = T) %>%
  dplyr::select(station, datetime, thermo.depth, top, bottom, secchi_m) %>%
  mutate(datetime = as.Date(datetime))

sonde.env <- bind_rows(sonde.epi, sonde.sed) %>%
  dplyr::select(station, date, parameter, result) %>%
  spread(parameter, result) %>%
  mutate(date = as.Date(date))

```

<EM: chemistry>

```{r}
chem.dat <- read.csv("./data/ConestogoLake_Chemistry_2017-2018.csv", header = T) %>%
  mutate(date = as.Date(date)) %>%
  filter(date < as.Date("2017-12-31"))

parms.list <- unique(chem.dat$parmUnit)[-c(2,13,16,17,19,20,24,25)]
```


```{r}

chem.epi <- chem.dat %>% filter(depth_m == 2, parmUnit %in% parms.list) %>%
  mutate(parameter = paste0(parmUnit,"_epi")) %>%
  dplyr::select(station, date, parameter, result)

chem.sed <- chem.dat %>% 
  filter(parmUnit %in% parms.list) %>%
  group_by(station, date, parmUnit) %>% 
  slice(n()) %>%
  mutate(parameter = paste0(parmUnit,"_sed")) %>%
  ungroup() %>%
  dplyr::select(station, date, parameter, result)

chem.env <- bind_rows(chem.epi, chem.sed) %>%
  dplyr::select(station, date, parameter, result) %>%
  spread(parameter, result)
```


```{r}
env <- chem.env %>%
  left_join(par.data, by = c("station"="site","date")) %>%
  left_join(sonde.env,by = c("station","date")) %>%
  left_join(lake.phys, by = c("station","date" = "datetime")) %>%
  mutate(DOY = yday(date), site.code = paste0(station, DOY)) %>%
  filter(station != "clw")
```

```{r}
env.mat <- read.csv("./data/envmatrix.csv") 
rownames(env.mat) <- env.mat$X
env.mat <- env.mat[,-1]

# standardize environmental matrix
# given the diversity of measurements in the environmental matrix, we need
# to standardize

## Standardization of all environmental variables
# Center and scale = standardize the variables (z-scores)
env.z <- decostand(env.mat, "standardize")
apply(env.z, 2, mean) # means = 0
apply(env.z, 2, sd) # standard deviations = 1
# Same standardization using the scale() function (which returns
# a matrix)
env.z <- as.data.frame(scale(env.mat))
```


```{r}

sig.parms <-
  c(
    "SRP_ug.L_sed",
    "TDP_ug.L_epi",
    "pH_avg_sed",
    "temp_C_avg_epi",
    "PAR",
    "SRP_ug.L_epi",
    "DO_mg.L_avg_epi",
    "TP_ug.L_epi",
    "pH_avg_epi",
    "temp_C_avg_sed"
  )
```

```{r loadPhytoMatrix}
spe.h <- read.csv("./data/phytomatrix_transformed.csv")
rownames(spe.h) <- spe.h$X
spe.h <- spe.h[,-1]

```

```{r RDA}

constraints <- env %>% select(station, DOY) %>% filter(DOY >= 186)
con.spe.rda <- rda(spe.h ~ DOY + station, constraints)
anova.cca(con.spe.rda, permutations = how(nperm = 999))

con.spe.rda <- rda(spe.h ~ ., env.z %>% select(sig.parms))
anova.cca(con.spe.rda, permutations = how(nperm = 999))

```


### cyanobacteria/toxin correlation

```{r, eval = T}
cyano.biomass.mat <- phyto.merge %>%
  filter(depth == 2, Phylum == "Cyanophyta", 
         Family != ".", Description.Code == "PHYT", station != "clw") %>% # select epi species level cyanos
  mutate(samp.no = paste0(station, DOY)) %>%
  dplyr::select(samp.no, species.code, biomass.mg.m3) %>%
  spread(species.code, biomass.mg.m3, fill = 0) %>% # create matrix of biomass data, fill NA's with 0
  dplyr::select(samp.no,P1041,P1076,P1133, P1075)

# only species that were visible in sample were counted. assume true 0's for everything else

toxins.mat <- tox.dat %>%
  filter(site != "clw") %>%
  mutate(samp.no = paste0(site, DOY)) %>%
  dplyr::select(samp.no, toxin, concentration) %>%
  spread(toxin, concentration, fill = 0)

# the toxin matrix has an extra date. Remove from matrix
toxins.mat <- toxins.mat[-1,]

merge.bio <- cyano.biomass.mat %>%
  left_join(toxins.mat, by = "samp.no") %>%
  mutate(site = as.factor(substr(samp.no, start = 1, stop = 3)),
         DOY = as.numeric(substr(samp.no, start = 4, stop = 6)))

pairs(merge.bio[,-c(1,8:9)])

#rownames(cyano.biomass.mat) <- cyano.biomass.mat


library(psych)
res <- corr.test(cyano.biomass.mat[,-1], toxins.mat[,-1], method = "kendall")
res

pander(res$p)

```

```{r RMANOVA-cyanobiomass-bygroup}

## did phytoplankton biomass change through time between sites?

cyano.stats <- phyto.com %>% #filter(Phylum == "Cyanophyta") %>% 
  filter(station != "clw") %>% 
  droplevels() %>%
  left_join(env.mat, by = c("station","DOY"))

cyano.stats <- as.data.frame(cyano.stats)

model <- lme(data = cyano.stats, 
           total.biomass ~ group * DOY * station,  
           random = ~1|group/station, 
           correlation = corAR1(form = ~ DOY|group/station), 
           method = "REML", na.action = na.omit)

model.fixed <- gls(total.biomass ~ group * DOY * station,
    correlation = corAR1(form = ~ DOY|group/station),
    data = cyano.stats,
    method = "REML")

anova(model, model.fixed) ## There is no benefit to having site as a random variable.
anova(model.fixed)

# calculate the least square mean
marginal = lsmeans(model.fixed,
                   ~ group/station:DOY)

cld(marginal,
    alpha   = 0.05,
    Letters = letters,     ### Use lower-case letters for .group
    adjust  = "tukey")     ###  Tukey-adjusted comparisons


# There is no difference between the sites through time.

# Check the model fit
histogram(residuals(model.fixed))

ACF(model.fixed,
    form = ~ DOY | Site)

```


```{r CYANOONLY-RMANOVA}
cyano.stats <- phyto.com %>% filter(Phylum == "Cyanophyta", station != "clw", DOY > 180) %>% 
  droplevels() %>%
  left_join(env.mat, by = c("station","DOY"))
cyano.stats <- as.data.frame(cyano.stats)

fit.ar1 <- lme(data = cyano.stats, 
           total.biomass ~ DOY * SRP_ug.L_epi * PAR * TDP_ug.L_epi *  temp_C_avg_epi,  
           random = ~ 1|station, 
           correlation = corAR1(form = ~ DOY|station, value = 0), 
           method = "REML", na.action=na.omit)

fit.arma <- lme(data = cyano.stats, 
           total.biomass ~ DOY,  
           random = ~1|station, 
           correlation = corARMA(form = ~ DOY|station, p = 1, q = 1), 
           method = "REML", na.action=na.omit)

fit.arma2 <- lme(data = cyano.stats, 
           total.biomass ~ DOY,  
           random = ~1|station, 
           correlation = corARMA(form = ~ DOY|station, p = 1, q = 2), 
           method = "REML", na.action=na.omit)


model <- lme(data = cyano.stats, 
           total.biomass ~  DOY * SRP_ug.L_epi * temp_C_avg_epi,  
           random = ~1|station, 
           correlation = corAR1(form = ~ DOY|station), 
           method = "REML", na.action = na.omit)

model.fixed <- gls(total.biomass ~ DOY * SRP_ug.L_epi * temp_C_avg_epi,
    correlation = corAR1(form = ~ DOY|station),
    data = cyano.stats,
    method = "REML")

model.fixed <- gls(total.biomass ~ DOY * station,
    correlation = corAR1(form = ~ DOY|station),
    data = cyano.stats,
    method = "REML")

anova(model, model.fixed) ## There is no benefit to having site as a random variable.
anova(model.fixed)

# calculate the least square mean
marginal = lsmeans(model.fixed,
                   ~ station:DOY)

cld(marginal,
    alpha   = 0.05,
    Letters = letters,     ### Use lower-case letters for .group
    adjust  = "tukey")     ###  Tukey-adjusted comparisons


# There is no difference between the sites through time.

# Check the model fit
histogram(residuals(model.fixed))

ACF(model.fixed,
    form = ~ DOY | Site)


```


```{r toxins-RMANOVA}
tox.raw <- read.csv("./data/Conestogo_toxins-LOQ-MOD.csv") %>%
  mutate(Sampling.date = as.Date(Sampling.date, format = "%Y-%m-%d"), 
         DOY = yday(Sampling.date)) 

tox.dat <- tox.raw %>%
  dplyr::select(site, Sampling.date, DOY, Depth..m., MC.LR, MC.RR, MC.YR, AP.A) %>%
  filter(site != "clw") %>%
  gather(key = toxin, value = concentration, MC.LR:AP.A) %>%
  mutate(concentration = concentration/1000)

tox.dat$concentration[is.na(tox.dat$concentration)] <-0
tox.dat$toxin <- as.factor(tox.dat$toxin)

# test model with random effects vs. just fixed effects
model <- lme(data = tox.dat, 
           concentration ~ toxin * DOY * site,  
           random = ~1|toxin/site, 
           correlation = corAR1(form = ~ DOY|toxin/site), 
           method = "REML", na.action=na.omit)

model.fixed <- gls(concentration ~ toxin * DOY * site,
    correlation = corAR1(form = ~ DOY|toxin/site),
    data = tox.dat,
    method = "REML")

anova(model, model.fixed) ## There is no benefit to having site as a random variable.
anova(model.fixed)

# calculate the least square mean
marginal = lsmeans(model.fixed,
                   ~ toxin/site:DOY)

cld(marginal,
    alpha   = 0.05,
    Letters = letters,     ### Use lower-case letters for .group
    adjust  = "tukey")     ###  Tukey-adjusted comparisons


# There is no difference between the sites through time.

# Check the model fit
histogram(residuals(model.fixed))

ACF(model.fixed,
    form = ~ DOY | Site)
```

```{r toxin-RMANOVA-table}
pander(anova(model.fixed))
```


